Consistency and Fault Tolerance Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project

### 2014-03-11 paper.tex
completed draft. Needs to be edited down for submission to 6 pages

\title{Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project}

@@

\begin{abstract}
With phase 1 of the Fast Forward Storage and IO Stack project complete, it is
an excellent opportunity to evaluate many of the decisions made to feed into
the phase 2 effort. The initial effort to define a next generation filesystem
has made admirable contributions in architecture and design. Formalizing the
general idea of data staging as burst buffers for the file system will help
manage the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a tranasctional mechanism to
manage faults and data visilibilty helps enable effective analytics without
having to work around the IO stack semantics.

@@

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
acitivies, such as two-phase, data sieving IO's data rearrangement process, can
be moved to the IOD layer offloading the communication load from the compute
nodes. IOD has three main purposes. First, if the optional burst buffer is
availble, it works as a fast cache absorbing write write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that should prevent a data set from being used. Third,
data processing operations can be placed in the IOD. This operations are
inteded to offer data rearrangement, filtering, and similar operations prior to
it reaching the central storage array.

@@

%Incorporated into the design of these layers and the function of some of these
%layers themsleves, such as the IOD layer, are features to manage both burst IO
%performance and faults. In particular, the IOD layer's use of burst buffers
%attempts to absorb a massive data dump from the compute area that is then
%trickled to the persistent storage area only if desired. Otherwise, it can be
%stored locally on optional SSDs for use by analysis routines eliminating the
%need to use slower rotational media as the intermediary between a simulation
%and the analysis code.

@@

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incoporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:optimizations} discusses some
of the optimizations offered by the current design and how and when these
optimizations will work best. It also explores some of the dark corners that
may yield performance issues with the current design. 
Section~\ref{sec:summary} disucsses the system overall with recommendations on
what design elements should be considered based on broader issues with current
HPC data centers. The paper is conclued in Section~\ref{sec:conclusion} with
a summary of the broad issues covered in the paper.

@@

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA paper in particular examines the potential costs and
advantages of where to place operators similar to the IOD proposed function
shipping. There are two key takeaways from that work. First, placement matters.
Depending on the communication intensity vs. computation intensity, where
along the IO path to place the operation can matter significantly. Second, and
more importantly, the amount of time spent processing for the operators was
stretched to the point where it consumed nearly all of the time between IO
operations. The given ratios of compute processes to staging process examined
is representative for future extreme scale platforms. If anything, the ratios
offer more staging processes than IOD processes would be available.

By concentrating this functionality into the storage stack, three problems
arise.  First, the amount of network bandwidth, IO bandwidth, and compute power
consumed for example operations from a single application is likely to
completely monopolize the IOD processes. Second, if space and time partitioning
is used instead, the functionality risks being too small to be useful. Third,
the hardware performance advantage for SSDs is questionable. Current NAND-based
flash devices top out at around 400 MB/sec. The key spec that is missing from
this number is that 400 MB/sec is a measure of the fixed number of available
IOPS multipled by the block size. This represents the ideal streaming
performance possible. The problem is that it costs an IOP to read 1 byte or 1
block (4KB or 8KB, depending on the device). It costs 1 IOP to write a full
block--usually. In some cases, it will cost 2 IOPs.  This accounts for the
required pre-erase write prior to writing to a reused block. In the worst case,
it can be 3 IOPS per write. This would be for a partial block write (read the
old block, erase, write the modified block). One-third of 400 MB/sec, about 133
MB/sec, is well below the streaming performance of HDDs.  Granted, there is
still rotational and seek latency to deal with for HDDs, but the advantage for
SSDs has evaporated and potentially turned into a penalty at a considerable
cost premium.  There are faster SSD solutions on the market that incorporate
DRAM for caching and using the PCIe bus, for example, but their price precludes
them from use in an extreme scale platform.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimatey relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

@@

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container serves to host a transaction and contains a collection of
objects. Conceptually, containers corresponds to a file in a traditional file
system. The objects in each container represent different data within a file.
The three initially defined object types are key-value store, array shard,
and blob. The easiest way to understand these types is to evaluate these from
the perspective of an HDF-5 file, the initial user interface layer. The
key-value store represents a collection of attributes. The array shard
represents part of a potentially multi-dimensional array. It is refered to as a
shard because it is likely a small piece of a globally defined array. The blob
represents a byte stream of arbitrary contents.  The fundamental difference
between an array shard and a blob is that the array shard has metadata
identifying its portion within the global, logical space while the blob is
simply a 1-dimensional array of bytes that is not shared across IOD nodes.
Should an operation be deployed into IOD to manipulate an array within a
container, it would operate on the array shards rather than on blobs. Given
this context, the transactions come in two forms.

@@

\subsection{Metadata Management}
While a central design goal of the Fast Forward project to eliminate the
requireent for a metadata server as a core component, it was not fully
accomplished.  This idea is not new and has been explored as part of the Light
Weight File Systems project~\cite{oldfield:lwfs}. In LWFS, the metadata service
is explicitly limited to a user task with the storage layer limited to data
storage/retrieval, authorization, and authentication. This approach proved
workable. For the Fast Forward project, while a similar goal is desired, it is
not completely accomplished.

The purpose behind the goal is to eliminate the serialized bottleneck caused by
having a centralized metadata service. Much of the design of IOD with objects
and continers does work to eliminate these bottlenecks.  However, IOD does not
go fully to a no metadata service model. Instead, the IOD layer describes a
key-value metadata service it provides. This service maintains four different
items:

@@

While a pure no metadata model would be ideal for performance, this current
proposal has a few challenges. By providing this basic functionality, metadata
management is split between the user layer and the IOD layer. Consdier the
following items. First, there are no list of containers. This is closer to the
no metadata ideal, but given the rest of the functionality, it exports to the
user level a requirement to track what containers are in the file system.
Second, given the flattening that occurs between IOD and DAOS, the mapping from
IOD objects to DAOS objects is unlikely to be 1-to-1. This is likely a list
instead. Further, given the ability to re-stripe in the IOD layer, the list is
not a list of objects completely included in the IOD object, but instead
objects from which part of the data came from some listed DAOS object.  Third,
sharding IOD objects makes sense based on the compute process connection a
shard has with the complete object.  Striping of an IOD object makes sense in
connection with the DAOS object(s) underneath, but striping within the IOD
layer is not discussed. Further, using striping from clients into the IOD layer
introduces a level of coordination that IOD seeks to avoid.  Fourth, given the
previous items, there is ambiguity with the maximum valid offset of the object.
This could be for the shard or for the collection of shard objects that
represent a distributed variable. Ideally, either a robust metadata management
system should be incorporated into the system, preferably separate from the IOD
layer or all of this functionality should be left to the user layer. The only
potential challenge is determining the container for a file open command and
what the transaction ID should be. This limited operation should be a
straightforward search using a hash or similar fast mechanism for examining a
list of container names and/or object IDs.

The \DDT project created a simplistic model for a metadata service that has
different features. It is described below and has been published
previously~\cite{lofstead:2012:txn-metadata}. Some of the major differences
include the following. First, \DDTns's metadata service has a way to get a list
of the objects in the metadata service. Since this was intended as a test
system to examine \DDT performance, we did not put the effort into a making it
a robust system. While it does not address having a file equivalence like the
container concept, it does introduce a way to figure out what is available.
Stored along with this object description is a list of the global dimensions
for array objects. One observtion made as part of creating this service is that
this is insufficient for many engineering codes that use non-regular meshes.
Second, for each object, a list of the equivalent to shards are kept with a
link back to the master object and the offset and size for each dimension for
that shard. This approach is not scalable because of the explosion in the
process count. Using a striping approach like IOD does is superior. Third,
there is no mapping to another layer with different object layout and counts.

@@

NEED TO ENUMERATE SOME OF THE CLIENT-SIDE ONLY PROBLEMS

@@

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordiation model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}. The
breakdown of the number of participants in each role is shown in
Table~\ref{tab:scaling}. For comparison, consider the Number of
Sub-Coordinators equivalent to an IOD process. The Processes Per
Sub-Coordinator represents the number of clients that use a particular IOD. For
these tests, we maintained a balanced distribution and always used at least two
sub-coordinators to slow down the processing.

@@

At a a high level, both \DDT and the IOD transactions have the same high-level
design. In both cases, a hierarchical model is employed. In the case of \DDTns,
it is a purely client-side tree using semi-synchronous messaging. The messaging
itself, in the current implementation, uses asynchronous MPI messages. The
synchronous component comes from the timeout mechanism used to detect faults.
It forces a level of coordiation and synchronization for the protocol. For IOD,
it is a server-side tree and fully asynchronous. In both cases, there is a
master in charge of managing the transaction and a collection of workers that
aggregate into the master through second-level leaders. Beyond that, there are
some significant differences. Some of the different choices made by IOD raise
some possible concern.

@@

Fourth, in the multi-leader model, using a count of client connections to
determine if transaction is complete is problematic.  First, should this be in
an evnironment where lost or repeated network messages are possible, then the
count could be wrong inappropriately. This is a general problem and unlikely
in the current environment, but is a concern on less integrated platforms.
The idea of a count precludes any tracking of expected messages from sources
introducing a potential hole. Second, the aggregation of completion messages
are based on the local aggregation point reporting to the transaction
leader the state.  There is no ability to detect or recover from a failure of
this IOD process that would report to the transaction leader. Third, because a
simple count of participating processes is sent with the beginning of the
transaction, each process is strictly limited to a single output operation.
This precludes a process writing multiple array shards, for example. To put
this in concrete terms, it means that a file could only contain a single,
globally distributed variable (at the root level of HDF5) OR an attribute per
process. If this count actually represents a begin/end transaction pair
instead, it suffers from a lack of accountability for missing object (shard)
writes. All it would indicate is that a process started and ended a transaction
connection. There is no validation that a process has done everything required
of it.

@@

\subsection{Comparision to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guarantee for when consensus will be achieved
without using slower synchronous calls. For the tight timing we wish to
support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments.

@@

\section{Optimizations}
\label{sec:optimizations}

THIS MUST BE RECONCILED WITH THE EARLIER TEXT THAT SAYS LARGELY THE SAME THING.

The ability to place functionality at the IOD layer was explored earlier. In
particular, PreDatA~\cite{zheng:2010:predata} demonstrated the advantage to
the approach as well as identified that the placement of the operators should
consider the compute nodes, the staging or burst buffer nodes, or offline
depending on the characteristics of the operation. This consideration should
be incorporated into the design. In particular, one of the key observations
is that the reduced process count may require the entire time between output
operations to perform many of the data processing tasks requested. This was
part of the motivation prompting evaluating placement decisions at the compute
node, in a staging area, or offline. In the case of the IOD design, it forces
a staging area deployment and then shares that staging area across all users
simultaneously. This is unlikely to be useful because of the limited compute
and communication capacity to spare to perform these operations at a bottleneck
in the IO path. The use of a separate staging area to perform these computations
intentionally selects a location separate from the IO path to avoid these
bottlenecks and potential scalability issues.

@@

THiS SECTION NEEDS SOME HELP

Consider a shared file system across an HPC data center. The current design
maintains the metadata in the IOD layer localized to a single machine
effectively making the data inaccessible from another platform.

There is a change in definitions between the IOD layer and the DAOS layer.
For the IOD layer, a container is a collection of objects. For the DAOS layer,
a continer is a collection of shards. For the IOD an object may be a shard of
a global array. For DAOS, a shard can host a set of DAOS objects. Having the
same names with locally correct, globally conflicting definitions serves to
confuse how the system should work.

@@

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the concensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our undertanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

We hope that the efforts made in the \DDT project and other efforts to explore
the requirements for this space, along with the analysis presented in this
paper will prove useful for the next phase of the Fast Forward project.

>>>>

\title{Consistency and Fault Tolerance Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project}

@@

\begin{abstract}
With phase 1 of the Fast Forward Storage and IO Stack project complete, it is
an excellent opportunity to evaluate many of the decisions made to feed into
the phase 2 effort. The initial effort to define a next generation file system
has made admirable contributions in architecture and design. Formalizing the
general idea of data staging as burst buffers for the file system will help
manage the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a transactional mechanism to
manage faults and data visibility helps enable effective analytics without
having to work around the IO stack semantics.

@@

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as two-phase, data sieving IO's data rearrangement process,
can be moved to the IOD layer offloading the communication load from the
compute nodes. IOD has three main purposes. First, if the optional burst buffer
is available, it works as a fast cache absorbing write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that would prevent a data set from being used. Third, data
processing operations can be placed in the IOD. These operations are intended
to offer data rearrangement, filtering, and similar operations prior to it
reaching the central storage array.

@@

%Incorporated into the design of these layers and the function of some of these
%layers themselves, such as the IOD layer, are features to manage both burst IO
%performance and faults. In particular, the IOD layer's use of burst buffers
%attempts to absorb a massive data dump from the compute area that is then
%trickled to the persistent storage area only if desired. Otherwise, it can be
%stored locally on optional SSDs for use by analysis routines eliminating the
%need to use slower rotational media as the intermediary between a simulation
%and the analysis code.

@@

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incorporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:summary} discusses the system
overall with recommendations on what design elements should be considered based
on broader issues with current HPC data centers. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
potential costs and advantages of where to place operators similar to the IOD
proposed function shipping. There are two key takeaways from that work. First,
placement matters.  Depending on the communication intensity vs. computation
intensity, where along the IO path to place the operation can matter
significantly. Second, and more importantly, the amount of time spent
processing for the operators was stretched to the point where it consumed
nearly all of the time between IO operations. The given ratios of compute
processes to staging process examined is representative for future extreme
scale platforms. If anything, the ratios offer more staging processes than IOD
processes would be available.

In the case of the IOD design, it forces a staging area deployment and then
shares the staging area across all users simultaneously. This is unlikely to be
useful because of the limited compute and communication capacity to spare to
perform these operations at a bottleneck in the IO path. The use of a separate
data staging area intentionally separate from the IO path allows using
operators on limited resources leaving the IO path clear for strictly data
movement.

By concentrating this functionality into the storage stack, three problems
arise.  First, the amount of network bandwidth, IO bandwidth, and compute power
consumed for example operations from a single application is likely to
completely monopolize the IOD processes. Second, if space and time partitioning
is used instead, the functionality risks being too small to be useful. Third,
the hardware performance advantage for SSDs is questionable. Current NAND-based
flash devices top out at around 400 MB/sec. The key spec that is missing from
this number is that 400 MB/sec is a measure of the fixed number of available
IOPS multiplied by the block size. This represents the ideal streaming
performance possible. The problem is that it costs an IOP to read 1 byte or 1
block (4KB or 8KB, depending on the device). It costs 1 IOP to write a full
block--usually. In some cases, it will cost 2 IOPs.  This accounts for the
required pre-erase write prior to writing to a reused block. In the worst case,
it can be 3 IOPS per write. This would be for a partial block write (read the
old block, erase, write the modified block). One-third of 400 MB/sec, about 133
MB/sec, is well below the streaming performance of HDDs.  Granted, there is
still rotational and seek latency to deal with for HDDs, but the advantage for
SSDs has evaporated and potentially turned into a penalty at a considerable
cost premium.  There are faster SSD solutions on the market that incorporate
DRAM for caching and using the PCIe bus, for example, but their price precludes
them from use in an extreme scale platform.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimately relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

@@

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container serves to host a transaction and contains a collection of
objects. Conceptually, containers corresponds to a file in a traditional file
system. The objects in each container represent different data within a file.
The three initially defined object types are key-value store, array shard,
and blob. The easiest way to understand these types is to evaluate these from
the perspective of an HDF-5 file, the initial user interface layer. The
key-value store represents a collection of attributes. The array shard
represents part of a potentially multi-dimensional array. It is referred to as
a shard because it is likely a small piece of a globally defined array. The
blob represents a byte stream of arbitrary contents.  The fundamental
difference between an array shard and a blob is that the array shard has
metadata identifying its portion within the global, logical space while the
blob is simply a 1-dimensional array of bytes that is not shared across IOD
nodes.  Should an operation be deployed into IOD to manipulate an array within
a container, it would operate on the array shards rather than on blobs. Given
this context, the transactions come in two forms.

@@

\subsection{Metadata Management}
While a central design goal of the Fast Forward project to eliminate the
requirement for a metadata server as a core component, it was not fully
accomplished.  This idea is not new and has been explored as part of the Light
Weight File Systems project~\cite{oldfield:lwfs}. In LWFS, the metadata service
is explicitly limited to a user task with the storage layer limited to data
storage/retrieval, authorization, and authentication. This approach proved
workable. For the Fast Forward project, while a similar goal is desired, it is
not completely accomplished.

The purpose behind the goal is to eliminate the serialized bottleneck caused by
having a centralized metadata service. Much of the design of IOD with objects
and containers does work to eliminate these bottlenecks.  However, IOD does not
go fully to a no metadata service model. Instead, the IOD layer describes a
key-value metadata service it provides. This service maintains four different
items:

@@

While a pure no metadata model would be ideal for performance, this current
proposal has a few challenges. By providing this basic functionality, metadata
management is split between the user layer and the IOD layer. Consider the
following items. First, there is no list of containers. This is closer to the
no metadata ideal, but given the rest of the functionality, it exports to the
user level a requirement to track what containers are in the file system.
Second, given the flattening that occurs between IOD and DAOS, the mapping from
IOD objects to DAOS objects is unlikely to be 1-to-1. This is likely a list
instead. Further, given the ability to re-stripe in the IOD layer, the list is
not a list of objects completely included in the IOD object, but instead
objects from which part of the data came from some listed DAOS object.  Third,
sharding IOD objects makes sense based on the compute process connection a
shard has with the complete object.  Striping of an IOD object makes sense in
connection with the DAOS object(s) underneath, but striping within the IOD
layer is not discussed. Further, using striping from clients into the IOD layer
introduces a level of coordination that IOD seeks to avoid.  Fourth, given the
previous items, there is ambiguity with the maximum valid offset of the object.
This could be for the shard or for the collection of shard objects that
represent a distributed variable. Ideally, either a robust metadata management
system should be incorporated into the system, preferably separate from the IOD
layer or all of this functionality should be left to the user layer. The only
potential challenge is determining the container for a file open command and
what the transaction ID should be. This limited operation should be a
straightforward search using a hash or similar fast mechanism for examining a
list of container names and/or object IDs.

The \DDT project created a simplistic model for a metadata service that has
different features. It is described below and has been published
previously~\cite{lofstead:2012:txn-metadata}. Some of the major differences
include the following. First, \DDTns's metadata service has a way to get a list
of the objects in the metadata service. Since this was intended as a test
system to examine \DDT performance, we did not put the effort into a making it
a robust system. While it does not address having a file equivalence like the
container concept, it does introduce a way to figure out what is available.
Stored along with this object description is a list of the global dimensions
for array objects. One observation made as part of creating this service is that
this is insufficient for many engineering codes that use non-regular meshes.
Second, for each object, a list of the equivalent to shards are kept with a
link back to the master object and the offset and size for each dimension for
that shard. This approach is not scalable because of the explosion in the
process count. Using a striping approach like IOD does is superior. Third,
there is no mapping to another layer with different object layout and counts.

@@

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks introduced.  First, a client could
create an entry in the metadata service that does not correspond to any objects
in the object store. Second, a client could create objects that have no
associated metadata entries. Third, updates to the metadata or object store
service should be an atomic operation, but due to a lack of coordination, a
window where the system is inconsistent appears.  Ultimately, the consistency
semantics required must be determined. If a metadata service is required and it
must be in sync with the object storage service, then additional work must be
performed. In traditional file systems, the metadata and object storage updates
are atomic. With decoupling metadata from object storage, should this atomicity
still be desired, it requires both the ability for the services to participate
in a task that is part of a larger atomic operation and a higher-level
mechanism to manage the atomic operation. Overall, while additional work is
required to maintain a client-side only metadata service, eliminating any
potential bottlenecks related to updating metadata related to the object
storage are eliminated. The burden of tracking striping and other metadata that
has traditionally been part of the metadata associated with the file system
will have to be maintained by the object storage service. The lack of a
centralized, serialized bottleneck to store that information improves
concurrency.

@@

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}. The
breakdown of the number of participants in each role is shown in
Table~\ref{tab:scaling}. For comparison, consider the Number of
Sub-Coordinators equivalent to an IOD process. The Processes Per
Sub-Coordinator represents the number of clients that use a particular IOD. For
these tests, we maintained a balanced distribution and always used at least two
sub-coordinators to slow down the processing.

@@

At a high level, both \DDT and the IOD transactions have the same high-level
design. In both cases, a hierarchical model is employed. In the case of \DDTns,
it is a purely client-side tree using semi-synchronous messaging. The messaging
itself, in the current implementation, uses asynchronous MPI messages. The
synchronous component comes from the timeout mechanism used to detect faults.
It forces a level of coordination and synchronization for the protocol. For
IOD, it is a server-side tree and fully asynchronous. In both cases, there is a
master in charge of managing the transaction and a collection of workers that
aggregate into the master through second-level leaders. Beyond that, there are
some significant differences. Some of the different choices made by IOD raise
some possible concern.

@@

Fourth, in the multi-leader model, using a count of client connections to
determine if transaction is complete is problematic.  First, should this be in
an environment where lost or repeated network messages are possible, then the
count could be wrong inappropriately. This is a general problem and unlikely
in the current environment, but is a concern on less integrated platforms.
The idea of a count precludes any tracking of expected messages from sources
introducing a potential hole. Second, the aggregation of completion messages
are based on the local aggregation point reporting to the transaction
leader the state.  There is no ability to detect or recover from a failure of
this IOD process that would report to the transaction leader. Third, because a
simple count of participating processes is sent with the beginning of the
transaction, each process is strictly limited to a single output operation.
This precludes a process writing multiple array shards, for example. To put
this in concrete terms, it means that a file could only contain a single,
globally distributed variable (at the root level of HDF5) OR an attribute per
process. If this count actually represents a begin/end transaction pair
instead, it suffers from a lack of accountability for missing object (shard)
writes. All it would indicate is that a process started and ended a transaction
connection. There is no validation that a process has done everything required
of it.

@@

\subsection{Comparison to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guarantee for when consensus will be achieved
without using slower synchronous calls. For the tight timing we wish to
support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments.

@@

@@

Consider a shared file system across an HPC data center. The current design
maintains the metadata in the IOD layer localized to a single machine
effectively making the data inaccessible from another platform. This requires
either the IOD layer be shared across machines, something clearly beyond the
design scope, or that an additional metadata service that extends the minimal
metadata service provided in the IOD layer. This higher level metadata service
would be required to be reachable from any machine that shares the file system.
As explained above, having this client-side only metadata service introduces
additional functionality requirements.

A point of confusion rather than a potential design challenge is the change in
definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
container is a collection of objects. For the DAOS layer, a container is a
collection of shards. For the IOD an object may be a shard of a global array.
For DAOS, a shard can host a set of DAOS objects. Having the same names with
locally correct, globally conflicting definitions serves to confuse how the
system should work.

@@

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the consensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our understanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

We hope that the efforts made in the \DDTns, Lightweight File Systems, and
other efforts to explore the requirements for this space, along with the
analysis presented in this paper will prove useful for the next phase of the
Fast Forward project.
